---
title: "ISLR Chapter 10 Unsupervised Learning"
author: "CAA"
date: "July 24, 2015"
output: html_document
---


R Computer Lab for Introduction to Statistical Learning, Chapter 10: Unsupervised Learning


# NCI60 Data Example

Cancer cell line microarray data: 6,830 gene expression measurements on 64 cancer cell lines.

### Examine data set:
```{r}
library(ISLR)
nci.labs = NCI60$labs
nci.data = NCI60$data
dim(nci.data)
nci.labs[1:4]
table(nci.labs)
```


### PCA on NCI60 Data
```{r}
pr.out = prcomp(nci.data, scale=TRUE)
```


### Plot the first few principal components:
```{r}
Cols = function(vec){
   cols = rainbow(length(unique(vec)))
   return (cols[as.numeric(as.factor(vec))])
}

par(mfrow=c(1,2))
plot(pr.out$x[,1:2], col=Cols(nci.labs), pch=19,
     xlab = "Z1", ylab="Z2")
plot(pr.out$x[,c(1,3)], col=Cols(nci.labs), pch=19, 
     xlab="Z1", ylab="Z2")
```

### Get summary:
```{r}
summary(pr.out)
plot(pr.out)
```


### Plot the PVE (Percent Variance Explained):
```{r}
pve = 100*pr.out$sdev^2 / sum(pr.out$sdev^2)
par(mfrow=c(1,2))
plot(pve, type="o", ylab="PVE", xlab="Principlal Component",
     col="blue")
plot(cumsum(pve), type="o", ylab="Cumulative PVE", 
     xlab="Principal Component", col="brown3")
```



### Clustering the observations of the NCI60 Data

```{r}
sd.data = scale(nci.data) # standardize data

par(mfrow=c(3,1))
data.dist = dist(sd.data)

plot(hclust(data.dist), labels=nci.labs, 
     main="Complete Linkage",xlab="",sub="",ylab="")


plot(hclust(data.dist, method="average"), labels=nci.labs,
     main="Average Linkage", xlab="", sub="",ylab="")


plot(hclust(data.dist, method="single"), labels=nci.labs,
     main="Single Linkage", xlab="", sub="",ylab="")

```



--------------- split above -------------

### Cut dendogram at height that yields 4 cluster:
```{r}
hc.out = hclust(dist(sd.data))
hc.clusters = cutree(hc.out,4)
table(hc.clusters,nci.labs)
```

### Plot the cut on the dendrogram:
```{r}
par(mfrow=c(1,1))
plot(hc.out,labels=nci.labs)
abline(h=139,col="red")

hc.out # summary of the object
```


### How does this compare to K-means?
```{r}
set.seed(2)
km.out = kmeans(sd.data,4,nstart=20)
km.clusters = km.out$cluster
table(km.clusters, hc.clusters)
```


### Hierarchical Clustering on PCA score vectors:
```{r}
hc.out = hclust(dist(pr.out$x[,1:5]))
plot(hc.out, labels=nci.labs, 
     main="Hier Clustering on First 5 Score Vectors")
table(cutree(hc.out,4), nci.labs)
```





